---
title: "R package"
author: "Jacob Courtney"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction



Congratulations for getting this far. To complete the evidence that you have made a working R package please answer the following by using the supplied rmd and upload the knitted html to CANVAS.

I will assume your package is called **MATH4753COUR0004** and you have packaged and documented functions `myci`, `myBoot`, `myncurve` and `myRandom`.

Your actual package name and functions will obviously be called something else.

# First function

Please place your first function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r first}
library(MATH4753COUR0004)
y=c(3,4,5,6,7)
confidence_interval=MATH4753COUR0004:::myci(y)
confidence_interval
```


# Second function

Please place your second function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r second}
library(MATH4753COUR0004)
set.seed(39)
sam=rnorm(25,mean=25,sd=10)
mb=MATH4753COUR0004:::myBoot(iter=10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="red")
mb
```


# Third function

Please place your third function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r third}
library(MATH4753COUR0004)
MATH4753COUR0004:::myncurve(mu=10,sigma=5,a=6)
```


# Forth function

Please place your forth function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r forth}
MATH4753COUR0004:::myRandom(1000,4,1) 
```

# The github repository



This is the address of my GITHUB repository:
<https://github.com/pojac3/MATH4753COUR0004.git>

Replace this with your own address.
